GEE models

### 2017-07-20 doc/manuscript.Rmd
mainly reformating paragraphs (added paris ref too)

Much of the previous experimental work on the role of NEFA in type 2 diabetes
utilized individual fatty acids such as palmitic acid (16:0) or oleic acid
(18:1n-9), or alternatively used specific oils such as soybean oil, which is
high in the polyunsaturated fatty acid (PUFA) linoleic acid (18:2n-6), as the
exposure to characterize the metabolic impact of total NEFA. However, fatty acids comprise
multiple molecules with diverse physiological functions, and few 
studies have analyzed the effects of a broader spectrum of fatty acids. Notably,
one study suggested that eicosapentaenoic acid (20:5n-3) can protect against the
lipotoxic effect of palmitic acid in the beta-cells [@Kato2008].

Despite a sizable literature studying the role of total NEFA concentration in 
diabetes, there are important gaps in this research. The majority of previous 
studies have used animal models or cell lines [@Maris2013a;@Deguil2011a], 
have been short term human trials 
[@Szendroedi2012a;@Daniele2014a;@Liang2013a;@Kehlenbrink2012a], or have
been epidemiological studies that only looked at total NEFA and not individual
fatty acids [@Miller2012a;@Johns2014a]. To date, there have
been no longitudinal studies examining the role of the composition of the serum NEFA
fraction on the pathogenesis of diabetes, which is critical given the protracted natural
history of diabetes and the growing appreciation of the divergent effects of
individual fatty acids. Therefore, our objective was to examine the 
association of serum NEFA composition on changes over time in insulin
sensitivity and beta-cell function in a longitudinal cohort. We hypothesized
that higher palmitic acid and lower PUFA such as
eicosapentaenoic acid in the NEFA fraction would associate with declining insulin sensitivity and
beta-cell function over 6 years.

>>>>

Much of the previous experimental work on the role of NEFA in type 2 diabetes 
utilized individual fatty acids such as palmitic acid (16:0) or oleic acid 
(18:1n-9), or alternatively used specific oils such as soybean oil, which is 
high in the polyunsaturated fatty acid (PUFA) linoleic acid (18:2n-6), as the 
exposure to characterize the metabolic impact of total NEFA. However, fatty
acids comprise multiple molecules with diverse physiological functions, and few 
studies have analyzed the effects of a broader spectrum of fatty acids. For instance,
one study suggested that eicosapentaenoic acid (20:5n-3) can protect against the
lipotoxic effect of palmitic acid in the beta-cells [@Kato2008].

Despite a sizable literature studying the role of total NEFA concentration in 
diabetes, there are important gaps in this research. The majority of previous 
studies have used animal models or cell lines [@Maris2013a;@Deguil2011a], have
been short term human trials 
[@Szendroedi2012a;@Daniele2014a;@Liang2013a;@Kehlenbrink2012a], or have been
epidemiological studies that only looked at total NEFA and not individual fatty
acids [@Miller2012a;@Johns2014a;@Charles1997a]. To date, there have been no longitudinal
studies examining the role of the composition of the serum NEFA fraction on the
pathogenesis of diabetes, which is critical given the protracted natural history
of diabetes and the growing appreciation of the divergent effects of individual
fatty acids. Therefore, our objective was to examine the association of serum
NEFA composition on changes over time in insulin sensitivity and beta-cell
function in a longitudinal cohort. We hypothesized that higher palmitic acid and
lower PUFA such as eicosapentaenoic acid in the NEFA fraction would associate
with declining insulin sensitivity and beta-cell function over 6 years.
